Inference using context-dependent models of mutation

### 2017-05-08 writeups/context-dependence.tex
.

1. Posterior coverage of parameters for Ising in 20 runs (more?).

2. With a CpG model with two 3mer motifs added, 1. fit single-base model; 2. add all 2mers; 3. add top one if significant by LR test and repeat; 4. do same for 3mers

@@

The usage of fast sparse matrix methods described above make likelihood computation fast enough
to use a Metropolis-Hasting-based Markov chain Monte Carlo method
to sample from the posterior distribution on parameters.
To test this,
we evolved $10^6$ sites from the Ising model (described above) with $\lambda=0.15$, $\beta=0.3$, and $\gamma=0.15$
for 1 unit of time.
We then placed an Exponential prior with mean 1 on both transitions $+ \to -$ and $- \to +$
(without constraining these to be equal)
and half-Gaussian priors with mean 0 and scale 3 on both $\beta$ and $\gamma$,
and 
We repeated this procedure separately on XX independently simulated sequences,
and show in Figure \ref{fig:ising_coverage}
the 95\% credible intervals for each parameter.

\begin{figure}
    \begin{center} 
    \end{center} 
    \caption{
        \label{fig:ising_coverage}}
\end{figure}

>>>>

@@

The usage of fast sparse matrix methods described above make likelihood computation fast enough
to use a Metropolis-Hasting-based Markov chain Monte Carlo method
to sample from the posterior distribution on parameters.
To test this,
we evolved $10^6$ sites from the Ising model (described above) with $\lambda=0.5$, $\beta=1.0$, and $\gamma=0.5$
for 1 unit of time, 
resulting in around 460,000 possible mutations
and differences at around 186,000 sites between initial and final sequence.
We then placed an Exponential prior with mean 1 on both transitions $+ \to -$ and $- \to +$
(without constraining these to be equal)
and half-Gaussian priors with mean 0 and scale 3 on both $\beta$ and $\gamma$,
counted $(5,3,1)$ T-mers,
and used the mcmc package \citep{mcmcR} to run a random walk sampler for 10,000 steps
(judged sufficient by examination of diagnostic plots).
We repeated this procedure separately on 60 independently simulated sequences,
and show the resulting credible intervals in Figure \ref{fig:ising_coverage}.
Poseterior medians are within half a percent of the truth for $+ \to -$ and $- \to +$ (corresponding to $\lambda$)
and the temperature ($\beta$), but showed around 2\% error for the magnetic field $\gamma$.
Credible intervals were somewhat too narrow, as the true value fell within 95\% credible intervals
for between 55\% and 90\% of the simulations;
which is likely due to remaining nonindependence at the window sizes used.

\begin{figure}
    \begin{center} 
        \includegraphics{writeup-plots/coverage_results}
    \end{center} 
    \caption{
        Credible intervals for the four parameters of the dynamic Ising model,
        relative to the true value,
        from 60 MCMC runs on independently simulated sequences of $10^6$ bases
        differing at around 18.6\% of the sites.
        \label{fig:ising_coverage}}
\end{figure}
